\textbf{Adaptation of the PHQ-4 Short Screening for Depression and Anxiety to increase its Sensitivity to Subclinical Variability}

### 2023-01-17 analysis.Rmd
update JAD

fig.width <- see::golden_ratio(7)
fig.height <- 7

@@

@@

data <- df |> 
  pivot_longer(everything(), names_to = "Item", values_to = "Answer") |> 
  group_by(Item, Answer) |> 
  summarise(n = n() / nrow(df)) |> 
  mutate(
    Facet = str_split_fixed(Item, "_", 2)[, 1],
    # Item = str_split_fixed(Item, "_", 2)[,2],
    Answer = case_when(
      Answer == 0 ~ "Not at all",
      Answer == 1 ~ "Once or twice",
      Answer == 2 ~ "Several days",
      Answer == 3 ~ "More than half the days",
      TRUE ~ "Nearly every day"
    ),
    Answer = fct_relevel(Answer, c("Not at all", "Once or twice", "Several days", "More than half the days",  "Nearly every day")),
    Item = case_when(
      Item == "Anxiety_1" ~ "A1: Feeling nervous, anxious or on edge",
      Item == "Anxiety_2" ~ "A2: Not being able to stop or control worrying",
      Item == "Depression_3" ~ "D1: Little interest or pleasure in doing things",
      TRUE ~ "D2: Feeling down, depressed, or hopeless"
    ))

p1 <- data |> 
  ggplot(aes(x = Answer, fill=Facet)) +
  geom_bar(aes(y = n), stat = "identity") +
  scale_y_continuous(labels=scales::percent, expand = c(0, 0)) +
  scale_fill_manual(values = c("Depression" = "#3F51B5", "Anxiety" = "#E91E63"), guide = "none") +
  labs(y = "Proportion of Answers", title = "A. Proportion of answers per item") +
  theme_modern(axis.title.space = 10) +
  theme(axis.text.x = element_text(angle = 45, hjust=1),
        axis.title.x = element_blank(),
        plot.title = element_text(face = "bold"),
        strip.background=element_rect(fill="#EEEEEE", colour="white")) +
  facet_wrap(~Item)
p1
```

@@

@@

```{r message=FALSE, warning=FALSE}
fig1 <- wrap_elements(p1) + wrap_elements(p2a / p2b + plot_annotation(title = "B. Information Curves per Response Type", theme = list(plot.title = element_text(face = "bold"))) + plot_layout(guides = "collect")) + plot_layout(widths = c(1, 1.5))

>>>>

fig.width <- see::golden_ratio(7)
fig.height <- 7
set.seed(333)

@@

```{r message=FALSE, warning=FALSE}
add_labels <- function(x) {
  x <- case_when(
      x == 0 ~ "Not at all",
      x == 1 ~ "Once or twice",
      x == 2 ~ "Several days",
      x == 3 ~ "More than half the days",
      TRUE ~ "Nearly every day"
    )
  fct_relevel(x, c("Not at all", "Once or twice", "Several days", "More than half the days",  "Nearly every day"))
}
```

@@

data <- df |> 
  pivot_longer(everything(), names_to = "Item", values_to = "Answer") |> 
  group_by(Item, Answer) |> 
  summarise(n = n() / nrow(df)) |> 
  mutate(
    Facet = str_split_fixed(Item, "_", 2)[, 1],
    # Item = str_split_fixed(Item, "_", 2)[,2],
    Answer = add_labels(Answer),
    Item = case_when(
      Item == "Anxiety_1" ~ "A1: Feeling nervous, anxious or on edge",
      Item == "Anxiety_2" ~ "A2: Not being able to stop or control worrying",
      Item == "Depression_3" ~ "D1: Little interest or pleasure in doing things",
      TRUE ~ "D2: Feeling down, depressed, or hopeless"
    ))

p1 <- data |> 
  ggplot(aes(x = Answer, fill=Facet)) +
  geom_bar(aes(y = n), stat = "identity") +
  scale_y_continuous(labels=scales::percent, expand = c(0, 0)) +
  scale_fill_manual(values = c("Depression" = "#3F51B5", "Anxiety" = "#E91E63"), guide = "none") +
  labs(y = "Proportion of Answers") +
  theme_modern(axis.title.space = 10) +
  theme(axis.text.x = element_text(angle = 45, hjust=1),
        axis.title.x = element_blank(),
        plot.title = element_text(face = "bold"),
        strip.text = element_text(size=rel(0.7)),
        strip.background=element_rect(fill="#EEEEEE", colour="white")) +
  facet_wrap(~Item)
p1 + ggtitle("Proportion of answers per item")
```

@@

```{r joint_prevalence, message=FALSE, warning=FALSE, fig.width=fig.width}
p1b <- rbind(
  data.frame(Scores = paste0(df$Anxiety_1, "_", df$Anxiety_2),
             Facet = "Anxiety"),
  data.frame(Scores = paste0(df$Depression_3, "_", df$Depression_4),
             Facet = "Depression")
) |> 
  group_by(Facet, Scores) |> 
  summarize(n = n() / nrow(df)) |> 
  separate(Scores, into = c("Q1", "Q2")) |> 
  mutate(Label = ifelse(Q1 == Q2, format_percent(n), "")) |> 
  mutate(Q1 = add_labels(as.numeric(Q1)),
         Q2 = add_labels(as.numeric(Q2))) |> 
  ggplot(aes(x = Q1, y = Q2)) +
  geom_tile(aes(fill = n)) +
  geom_text(aes(label = Label), size=2, color="white") +
  scale_y_discrete(expand = c(0, 0)) +
  scale_x_discrete(expand = c(0, 0)) +
  scale_fill_gradientn(colors=c("white", "#2196F3", "#3F51B5", "#673AB7"), labels = scales::percent) +
  labs(fill = "% Participants", x = "Item 1", y = "Item 2") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust=1),
        panel.grid.major = element_blank(),
        plot.title = element_text(face = "bold", hjust=0),
        strip.background=element_rect(fill="#EEEEEE", colour="white"),
        strip.text = element_text(face = "bold")) +
  facet_wrap(~Facet) 
p1b + ggtitle("Joint prevalence of responses")
```

@@

# Figures

```{r message=FALSE, warning=FALSE}
fig1 <- wrap_elements(wrap_elements(p1 + plot_annotation(title = "A. Proportion of answers per item", theme = list(plot.title = element_text(face = "bold")))) / 
                        wrap_elements(p1b + plot_annotation(title = "B. Joint prevalence of responses", theme = list(plot.title = element_text(face = "bold"))))) + 
  wrap_elements(p2a / p2b + plot_annotation(title = "C. Information curves per response type", theme = list(plot.title = element_text(face = "bold"))) + plot_layout(guides = "collect")) + plot_layout(widths = c(1, 1.5))

### 2023-01-17 paper/1_PA/manuscript.Rmd
update JAD

keywords          : "PHQ-4, depression, anxiety, brief questionnaire, short scale"
wordcount         : "1158"
bibliography      : references.bib
floatsintext      : yes
linenumbers       : yes
draft             : no
mask              : no
figurelist        : yes
tablelist         : no
footnotelist      : no
classoption       : "man"
output            : papaja::apa6_pdf
csl: utils/apa.csl
header-includes:
  - \usepackage[labelfont=bf, font={scriptsize, color=gray}]{caption}
editor_options:
  chunk_output_type: console
---

@@

knitr::opts_chunk$set(
  collapse = TRUE,
  dpi = 450,
  fig.width = see::golden_ratio(9),
  fig.height = 9,
  fig.path = "../figures/"
)

@@

The sample consists of 485 English-speaking participants (Mean age = 30.1, SD = 10.1, range: [18, 73]; Sex: 50.3\% females, 49.7\% males) from the general population who were administered the refined PHQ-4 online as part of another study, which data is available in open-access at https://github.com/RealityBending/IllusionGameReliability. This study was approved by the NTU Institutional Review Board (NTU IRB-2022-187). All participants provided their informed consent prior to participation and were incentivized after completing the study.

@@

```{r message=FALSE, warning=FALSE, fig.cap="A. Proportion of answers of each type. B. Item Information Curves from IRT showing the coverage by each item and response of the latent dimension. Typically, an optimally informative item would display a large coverage over theta, with each response presenting a narrow coverage (high discrimination between different levels).", out.width="100%"}
knitr::include_graphics("../figures/figure1.png")
```

@@

```{r message=FALSE, warning=FALSE, fig.cap="Normalized scoring of the items taking into account the non-linear spacing between responses (based on the peak of their measurement information).", out.width="100%"}
knitr::include_graphics("../figures/figure2.png")
```

@@

The dataset analysed during the current study are available in the GitHub repository https://github.com/DominiqueMakowski/PHQ4R. We have created an R function for easy scoring according to the two methods described, which can be used as follows:

@@

This work was supported by the Presidential Postdoctoral Fellowship Grant (NTU-PPF-2020-10014) from Nanyang Technological University (awarded to DM).

>>>>

keywords          : "PHQ-4; depression; anxiety; brief questionnaire; short scale"
wordcount         : "1158"
bibliography      : references.bib
floatsintext      : yes
linenumbers       : yes
draft             : no
mask              : yes
figurelist        : yes
tablelist         : no
footnotelist      : no
classoption       : "man"
output            : papaja::apa6_pdf
csl: "../utils/apa.csl"
header-includes:
  - \usepackage[labelfont=bf, font={scriptsize, color=gray}]{caption}
editor_options:
  chunk_output_type: console
---

@@

knitr::opts_chunk$set(
  collapse = TRUE,
  dpi = 450,
  fig.width = see::golden_ratio(9),
  fig.height = 9,
  fig.path = "../../figures/"
)

@@

The sample consists of 485 English-speaking participants (Mean age = 30.1, SD = 10.1, range: [18, 73]; Sex: 50.3\% females, 49.7\% males) from the general population who were administered the refined PHQ-4 online as part of another study, which data is available in open-access at https://github.com/RealityBending/IllusionGameReliability. This study was approved by the 
Institutional Review Board **[Masked for Review]**.
<!-- NTU Institutional Review Board (NTU IRB-2022-187).  -->
All participants provided their informed consent prior to participation and were incentivized after completing the study.

@@

```{r message=FALSE, warning=FALSE, fig.cap="A. Proportion of answers of each type. B. Item Information Curves from IRT showing the coverage by each item and response of the latent dimension. Typically, an optimally informative item would display a large coverage over theta, with each response presenting a narrow coverage (high discrimination between different levels).", out.width="100%"}
knitr::include_graphics("../../figures/figure1.png")
```

@@

```{r message=FALSE, warning=FALSE, fig.cap="Normalized scoring of the items taking into account the non-linear spacing between responses (based on the peak of their measurement information).", out.width="100%"}
knitr::include_graphics("../../figures/figure2.png")
```

@@

The dataset analysed during the current study are available in the GitHub repository **[Masked for Review].** 
<!-- https://github.com/DominiqueMakowski/PHQ4R. -->
We have created an R function for easy scoring according to the two methods described, which can be used as follows:

@@

This work was supported by the **[Masked for Review]**.
<!-- Presidential Postdoctoral Fellowship Grant (NTU-PPF-2020-10014) from Nanyang Technological University (awarded to DM). -->

### 2023-01-17 paper/1_PA/manuscript.tex
update JAD

  man,floatsintext]{apa6}

@@

\linenumbers
\usepackage{csquotes}
\usepackage[titles]{tocloft}
\cftpagenumbersoff{figure}
\renewcommand{\cftfigpresnum}{\itshape\figurename\enspace}
\renewcommand{\cftfigaftersnum}{.\space}
\setlength{\cftfigindent}{0pt}
\setlength{\cftafterloftitleskip}{0pt}
\settowidth{\cftfignumwidth}{Figure 10.\qquad}
\usepackage[labelfont=bf, font={scriptsize, color=gray}]{caption}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Adaptation of the PHQ-4 Short Screening for Depression and Anxiety to increase its Sensitivity to Subclinical Variability},
  pdfauthor={Dominique Makowski1, An Shu Te1, \& S.H. Annabel Chen1, 2, 3, 4},
  pdflang={en-EN},
  pdfkeywords={PHQ-4, depression, anxiety, brief questionnaire, short scale},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

The sample consists of 485 English-speaking participants (Mean age = 30.1, SD = 10.1, range: {[}18, 73{]}; Sex: 50.3\% females, 49.7\% males) from the general population who were administered the refined PHQ-4 online as part of another study, which data is available in open-access at \url{https://github.com/RealityBending/IllusionGameReliability}. This study was approved by the NTU Institutional Review Board (NTU IRB-2022-187). All participants provided their informed consent prior to participation and were incentivized after completing the study.

@@

\begin{figure}
\includegraphics[width=1\linewidth]{../figures/figure1} \caption{A. Proportion of answers of each type. B. Item Information Curves from IRT showing the coverage by each item and response of the latent dimension. Typically, an optimally informative item would display a large coverage over theta, with each response presenting a narrow coverage (high discrimination between different levels).}\label{fig:unnamed-chunk-2}
\end{figure}

@@

\begin{figure}
\includegraphics[width=1\linewidth]{../figures/figure2} \caption{Normalized scoring of the items taking into account the non-linear spacing between responses (based on the peak of their measurement information).}\label{fig:unnamed-chunk-3}
\end{figure}

@@

The dataset analysed during the current study are available in the GitHub repository \url{https://github.com/DominiqueMakowski/PHQ4R}. We have created an R function for easy scoring according to the two methods described, which can be used as follows:

@@

This work was supported by the Presidential Postdoctoral Fellowship Grant (NTU-PPF-2020-10014) from Nanyang Technological University (awarded to DM).

>>>>

  man,mask,floatsintext]{apa6}

@@

\linenumbers
\usepackage{csquotes}
\usepackage[titles]{tocloft}
\cftpagenumbersoff{figure}
\renewcommand{\cftfigpresnum}{\itshape\figurename\enspace}
\renewcommand{\cftfigaftersnum}{.\space}
\setlength{\cftfigindent}{0pt}
\setlength{\cftafterloftitleskip}{0pt}
\settowidth{\cftfignumwidth}{Figure 10.\qquad}
\usepackage[labelfont=bf, font={scriptsize, color=gray}]{caption}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Adaptation of the PHQ-4 Short Screening for Depression and Anxiety to increase its Sensitivity to Subclinical Variability},
  pdflang={en-EN},
  pdfkeywords={PHQ-4, depression, anxiety, brief questionnaire, short scale},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

@@

The sample consists of 485 English-speaking participants (Mean age = 30.1, SD = 10.1, range: {[}18, 73{]}; Sex: 50.3\% females, 49.7\% males) from the general population who were administered the refined PHQ-4 online as part of another study, which data is available in open-access at \url{https://github.com/RealityBending/IllusionGameReliability}. This study was approved by the
Institutional Review Board \textbf{{[}Masked for Review{]}}.
All participants provided their informed consent prior to participation and were incentivized after completing the study.

@@

\begin{figure}
\includegraphics[width=1\linewidth]{../../figures/figure1} \caption{A. Proportion of answers of each type. B. Item Information Curves from IRT showing the coverage by each item and response of the latent dimension. Typically, an optimally informative item would display a large coverage over theta, with each response presenting a narrow coverage (high discrimination between different levels).}\label{fig:unnamed-chunk-2}
\end{figure}

@@

\begin{figure}
\includegraphics[width=1\linewidth]{../../figures/figure2} \caption{Normalized scoring of the items taking into account the non-linear spacing between responses (based on the peak of their measurement information).}\label{fig:unnamed-chunk-3}
\end{figure}

@@

The dataset analysed during the current study are available in the GitHub repository \textbf{{[}Masked for Review{]}.}
We have created an R function for easy scoring according to the two methods described, which can be used as follows:

@@

This work was supported by the \textbf{{[}Masked for Review{]}}.

### 2023-01-17 paper/2_JAD/manuscript.Rmd
update JAD

  - id            : "1"
    institution   : "School of Social Sciences, Nanyang Technological University, Singapore"
  - id            : "2"
    institution   : "LKC Medicine, Nanyang Technological University, Singapore"
  - id            : "3"
    institution   : "National Institute of Education, Singapore"
  - id            : "4"
    institution   : "Centre for Research and Development in Learning, Nanyang Technological University, Singapore"
abstract: |
  The PHQ-4 is an ultra brief (4 items) screening questionnaire for depression and anxiety. We propose to add one aditional response option ("Once or twice", in between "Not at all" and "Several days") to improve its sensitivity to milder alterations, and thus increase its usefulness in subclinical populations. Using Item Response Theory (IRT), we show that this new option does indeed capture specific portions of the measured constructs, and provide a scoring procedure that preserve the compatability with past versions.

@@

keywords          : "PHQ-4; depression; anxiety; brief questionnaire; short scale"
wordcount         : "1158"
bibliography      : references.bib
floatsintext      : yes
linenumbers       : yes
draft             : no
mask              : yes
figurelist        : yes
tablelist         : no
footnotelist      : no
classoption       : "man"
output            : papaja::apa6_pdf
csl: "../utils/apa.csl"
header-includes:
  - \usepackage[labelfont=bf, font={scriptsize, color=gray}]{caption}
editor_options:
  chunk_output_type: console
---

@@

While the scale has been validated and used in the general population [@lowe20104; @hajek2020prevalence], its initial purpose was to reliably discriminate and identify potential MDD/GAD patients. This discriminative goal materializes in the scale's design and the existence of categorical cut-offs, which does not necessary entail a strong sensitivity to milder mood alterations. In particular, the gap between the two lowest possible answers, "Not at all" and "Several days", is quite large and leaves out the possibility of more subtle occurrences. While this is not necessarily an issue in clinical and diagnostic contexts, it might lead to a sub-optimal discrimination of affective levels on the lower end of the spectrum, important for instance in the context of subclinical variability quantifications. The goal of this study is hence to enhance, with minimal changes to the original scale, the sensitivity to mild mood level inflections.

@@

The sample consists of 485 English-speaking participants (Mean age = 30.1, SD = 10.1, range: [18, 73]; Sex: 50.3\% females, 49.7\% males) from the general population who were administered the refined PHQ-4 online as part of another study, which data is available in open-access at https://github.com/RealityBending/IllusionGameReliability. This study was approved by the 
Institutional Review Board **[Masked for Review]**.
<!-- NTU Institutional Review Board (NTU IRB-2022-187).  -->
All participants provided their informed consent prior to participation and were incentivized after completing the study.

@@

```{r message=FALSE, warning=FALSE, fig.cap="A. Proportion of answers of each type. B. Item Information Curves from IRT showing the coverage by each item and response of the latent dimension. Typically, an optimally informative item would display a large coverage over theta, with each response presenting a narrow coverage (high discrimination between different levels).", out.width="100%"}
knitr::include_graphics("../../figures/figure1.png")
```

@@

We propose two types of scoring procedures. The first aims at
<!-- In order to  -->
minimally disrupting the original scale and making its scores comparable, enabling comparisons across studies and the usage of developed cut-offs and norms. We suggest to score the new option, "Once or twice", as 0.5, occupying the middle space between 0 and 1 (preserving in this way the total score range of 0 - 8 per dimension).

@@

The objective of this study was to test the introduction of a "Once or twice" response option to the PHQ-4 to enhance its sensitivity to milder mood fluctuations. The fact that this new response option was the most prevalent response made by participants in our study is in itself evidence for its usefulness. IRT analysis further revealed that this response captures with precision a unique portion of anxiety and depression. Taken together, our results suggest that adding this option response increases the scale's potential to discriminate average mood levels (which are superior to zero) from lower-end extremes (the true zero).

The hypothesized added value in terms of finer sensitivity should be confirmed by cross-validating the PHQ-4R with more comprehensive measures of anxiety and depression. Additionally, evidence comparing the refined version against the original PHQ-4 would be useful to better understand whether the new "Once or twice" option pulls participants that would have otherwise answer "Not at all" or "Several days". Finally, future studies should further corroborate our statistical validation and test the usefulness of this revision on more targeted samples (e.g., clinical populations) to see if it affects the cut-offs proposed in the literature (and whether it allows for a better discrimination of mood disorders), and refine if need be the normalized scoring estimates in larger sample.

@@

The dataset analysed during the current study are available in the GitHub repository **[Masked for Review].** 
<!-- https://github.com/DominiqueMakowski/PHQ4R. -->
We have created an R function for easy scoring according to the two methods described, which can be used as follows:

@@

This work was supported by the **[Masked for Review]**.
<!-- Presidential Postdoctoral Fellowship Grant (NTU-PPF-2020-10014) from Nanyang Technological University (awarded to DM). -->

>>>>

  - id            : "1"
    institution   : "School of Social Sciences, Nanyang Technological University, Singapore"
  - id            : "2"
    institution   : "LKC Medicine, Nanyang Technological University, Singapore"
  - id            : "3"
    institution   : "National Institute of Education, Singapore"
  - id            : "4"
    institution   : "Centre for Research and Development in Learning, Nanyang Technological University, Singapore"
abstract: |
  The PHQ-4 is an ultra-brief (4 items) screening questionnaire for depression and anxiety. We propose to add one additional response option ("Once or twice", in between "Not at all" and "Several days") to improve its sensitivity to milder alterations, and thus increase its usefulness in subclinical populations. Using Item Response Theory (IRT), we provide evidence that this new option does indeed capture specific portions of the measured constructs in the general population.

@@

keywords          : "PHQ-4; depression; anxiety; brief questionnaire; short scale"
wordcount         : "1249"
bibliography      : references.bib
floatsintext      : yes
linenumbers       : yes
draft             : no
mask              : no
figurelist        : yes
tablelist         : no
footnotelist      : no
classoption       : "man"
output            : papaja::apa6_pdf
csl: "../utils/apa.csl"
header-includes:
  - \usepackage[labelfont=bf, font={scriptsize, color=gray}]{caption}
editor_options:
  chunk_output_type: console
---

@@

While the scale has been validated and used in the general population and non-clinical samples [@lowe20104; @hajek2020prevalence], its initial purpose was to reliably discriminate and identify potential MDD/GAD patients. This discriminative goal materializes in the scale's design and the existence of categorical cut-offs, which does not necessary entail a strong sensitivity to milder mood alterations. In particular, the gap between the two lowest possible answers, "Not at all" and "Several days", is quite large and leaves out the possibility of more subtle occurrences. While this is not necessarily an issue in clinical and diagnostic contexts, it might lead to a sub-optimal discrimination of affective levels on the lower end of the spectrum, important for instance in the context of subclinical variability quantifications. The goal of this study is hence to enhance, with minimal changes to the original scale, the sensitivity to mild mood level inflections.

@@

The sample consists of 485 English-speaking participants (Mean age = 30.1, SD = 10.1, range: [18, 73]; Sex: 50.3\% females, 49.7\% males) from the general population recruited via *Prolific*, a crowd-sourcing platform recognized for providing high quality data [@peer2022data]. The only inclusion criterion was a fluent proficiency in English to ensure that the task instructions would be well-understood. Note that the participants were administered the refined PHQ-4 online as part of another study, which data is available in open-access at https://github.com/RealityBending/IllusionGameReliability. This study was approved by the 
<!-- Institutional Review Board **[Masked for Review]**. -->
NTU Institutional Review Board (NTU IRB-2022-187).
All participants provided their informed consent prior to participation and were incentivized after completing the study.

@@

```{r message=FALSE, warning=FALSE, fig.cap="A) Proportion of answers of each type. B) Prevalence of answer pairs. C) Item Information Curves from IRT showing the coverage by each item and response of the latent dimension. Typically, an optimally informative item would display a large coverage over theta, with each response presenting a narrow coverage (high discrimination between different levels).", out.width="100%"}
knitr::include_graphics("../../figures/figure1.png")
```

@@

We propose two types of scoring procedures. The first aims at
minimally disrupting the original scale and making its scores comparable, enabling comparisons across studies and the usage of developed cut-offs and norms. We suggest to score the new option, "Once or twice", as 0.5, occupying the middle space between 0 and 1 (preserving in this way the total score range of 0 - 8 per dimension).

@@

The objective of this study was to test the introduction of a "Once or twice" response option to the PHQ-4 to enhance its sensitivity to milder mood fluctuations. The fact that this new response option was the most prevalent response made by participants in our study is in itself evidence for its usefulness. IRT analysis further revealed that this response captures with precision a unique portion of the variability in the latent factors measured by the instrument. Taken together, our results suggest that adding this option response increases the scale's potential to discriminate average mood levels (which are superior to zero) from lower-end extremes (the true zero).

It is important to emphasize that our study is a statistical validation in the general population, which analysed how different responses relate to one another. It is a first step providing supporting evidence for the proposed change. However, the structural gains that we observed in terms off finer sensitivity should be confirmed 1) by cross-validating the PHQ-4R with more comprehensive measures of anxiety and depression, and 2) testing the usefulness of this revision on more targeted samples (e.g., clinically characterized populations) to investigate the impact on existing cut-offs and on the detection of mood disorders. Additionally, evidence comparing the refined version against the original PHQ-4 would be useful to better understand whether the new "Once or twice" option pulls participants that would have otherwise answer "Not at all" or "Several days".
<!-- , and refine if need be the normalized scoring estimates in larger sample. -->

@@

The dataset analysed during the current study are available in the GitHub repository 
<!-- **[Masked for Review].**  -->
https://github.com/DominiqueMakowski/PHQ4R.
We have created an R function for easy scoring according to the two methods described, which can be used as follows:

@@

This work was supported by the 
<!-- **[Masked for Review]**. -->
Presidential Postdoctoral Fellowship Grant (NTU-PPF-2020-10014) from Nanyang Technological University (awarded to DM).
